---
title: "Bioconductor peeks at pangenomes, 2023"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Bioconductor peeks at pangenomes, 2023}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

State of the art work
on a human pangenome reference (as of 2023) is described
in [Liao et al.](https://www.nature.com/articles/s41586-023-05896-x)
The first draft "contains 47 phased, diploid assemblies from a cohort of genetically diverse individuals".
The purpose of this R package is to demonstrate how Bioconductor
data structures can be used to explore resources from the first draft.

# Minigraphs presented as GRanges

From the Liao et al. paper:

> Minigraph builds a pangenome by starting from a reference assembly, here 
GRCh38, and iteratively and progressively adds in additional assemblies, recording only 
SVs ≥ 50 bases. It admits complex variants, including duplications and inversions.

```{r setup,message=FALSE}
library(BiocHPRClk)
library(S4Vectors)
data(minigr_GRCh38)
minigr_GRCh38
```

There is a description of the columns in the table:
```{r lkmd}
cat(metadata(minigr_GRCh38)[[1]])
```
